TITLE by YOUR_NAME_HERE
Wine Quality by Naseer Faheem 

========================================================
```{r}
getwd()
#Mac
setwd("/Users/naseerfaheem/Desktop/google_drive/Workspace/Udacity/data-analyst/term2/07-EDA/final-project")
#Windows: 
setwd("D:/Google_Drive/Workspace/Udacity/data-analyst/term2/07-EDA/final-project")
getwd()
```

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(reshape2)
library(RColorBrewer)
library(GGally)

```
Questions: 
Alcohol content vs. quality
pH vs. quality 
density vs. quality. 
residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet
 chlorides: the amount of salt in the wine
    9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

   11 - alcohol: the percent alcohol content of the wine
```{r echo=FALSE, Load_the_Data}
# Load the Data
r_wine <- read.csv("wineQualityReds.csv")
r_wine
dim(r_wine)
str(r_wine)
w_wine <- read.csv("wineQualityWhites.csv")
w_wine
dim(w_wine)

#create a varaible to distinguish each dataset:
r_wine["type"] <- "red"
w_wine["type"] <- "white"

#combining both datasets into one: 
wine <- rbind(r_wine, w_wine)
#transforming the quality from integer to a factor
wine$quality <- factor(wine$quality, ordered = T)


str(wine)
names(wine)
length(names(wine))
head(wine)
tail(wine)
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

>This tidy data set contains 1,599 red wines with 12 variables and 4898 white wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent). For the sake of analysis, I combined both datasets into one dataset. I created a new variable designating the type of the wine, whether red or white wine. 

>Before going furthur in the analysis, its import to know the variables in the dataset and what they represent. The following two sections describe each variable in great detail so that it makes it easier to understand the result of the dataset. 

##Attribute information:
   For more information, read [Cortez et al., 2009].
   
   Input variables (based on physicochemical tests):
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)
   
##Description of attributes:
   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
   2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
   3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
   4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet
   5 - chlorides: the amount of salt in the wine
   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
   7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
   8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content
   9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
   11 - alcohol: the percent alcohol content of the wine
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)


# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
#Running the summary of the dataset to see everyting at a glance: 
summary(wine)
#One thing that standout is the risidual sugar. The mean of the variable is 5.44 whereas the max value is 65.8. It looks like we might have some outliers, which include very sweet wines. We could also notice that the mean wine quality  5.81 which we have minimums of 3 and maximums of 9. It shows that no wine recieved a quality of a perfect 10. Similarly, no wine recieved a quality of lower than 3. 
table(wine$residual.sugar)
# If we look at the distribution of the residual sugar, we can see that there is only wine with the residual sugar value of 65.8. 
```

# 1. Univariate Plotting: 
In this section, I will be plotting the distribution of the individual variable of the dataset to get a thourgh understanding. It will also help us see where the the majority of the observations are for each variable. After plotting all the variables individually, I will pick a few 

## 1.1.1 Fixed Acidity: 
```{r fixed acidity}
#To illustrate more, lets see a histogram of the residual sugar distribution: 
ggplot(wine, aes(fixed.acidity, fill=type))+
  geom_bar(binwidth = 1)+
  scale_x_continuous(breaks = seq(3, 16, 1), lim = c(3, 16))+
  scale_y_sqrt()+
  xlab("Wine Count")+
  ylab("Wine fixed acidity")

```
```{r fixed acidity summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(fixed.acidity), 
            median = median(fixed.acidity), 
            n = n())
```
>1.1.2 It looks like the fixed acidity for both wine types have a similar distribution. We could also see that the none of white wines have a fixed acidity of 13 or 15. We could also notice that the mean for fixed acidity of red is about 2g higher than the white wine. 

## 1.2.1 Volatile Acidity: 
```{r volatile.acidity}
#To illustrate more, lets see a histogram of the residual sugar distribution: 
ggplot(wine, aes(volatile.acidity, fill=type))+
  geom_bar(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 2, 0.2))+
  scale_y_sqrt()
```
```{r volatile acidity summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(volatile.acidity), 
            median = median(volatile.acidity), 
            n = n())
```
>1.2.2 Volatile acidity has almost a similar distribution for both wine types. It seems that red wine has a little higher volatile acitidy (0.53) whereas white volatile acidity level is low (0.28), which makes sense becuase a wine with too more acetic acid could lead to an unpleaseant vinegar taste. 


## 1.3.1 Citric Acid: 
```{r citric.acid}
ggplot(wine, aes(citric.acid, fill=type))+
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 2, 0.1))+
  scale_y_sqrt()
```
```{r citric acid summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(citric.acid), 
            median = median(citric.acid), 
            n = n())
```
> 1.3.2 Both wines appear to have normal distribution for Citric Acid levels, which is below 1. However, there are a few outliers for white wine which has a citric acid level of higher than 1.0.

## 1.4.1 Residual Sugar: 
```{r residual.sugar}
ggplot(wine, aes(residual.sugar, fill=type))+
  geom_histogram(binwidth = 2)+
  scale_x_continuous(breaks = seq(0, 66, 3))+
  scale_y_sqrt()
```

```{r residual sugar summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(residual.sugar), 
            median = median(residual.sugar), 
            n = n())
```
> 1.4.2 Both wine are have skewed distributions. Here, the white wine appears to have one outlier which has a residual sugar level of 66 grams/liter. It is pretty rare to find a wine with that much sugar in it. In general it looks like white wines higher sugar contents despite having an outlier. I base my assumption on the median of the white wine. 


## 1.5.1 Chlorides: 
```{r chlorides}
ggplot(wine, aes(chlorides, fill=type))+
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 2, 0.1))+
  scale_y_sqrt()
```

```{r chlorides summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(chlorides), 
            median = median(chlorides), 
            n = n())
```

>1.5.2 As we can see in both the plot and the summary table that red wines have higher levels of chlorides. It makes sense since we noticed earlier that white wine has higher residual sugar level. 


## 1.6.1 Free Sulfur Dioxide: 
```{r free.sulfur.dioxide}
ggplot(wine, aes(free.sulfur.dioxide, fill=type))+
  geom_histogram(binwidth = 25)+
  scale_x_continuous(breaks = seq(0, 400, 25))+
  scale_y_sqrt()

```

```{r free.sulfur.dioxide summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(free.sulfur.dioxide), 
            median = median(free.sulfur.dioxide), 
            n = n())
```
>1.6.2 The white wine has on average about 20 mg more free sulfur dioxide than the red wine. It will be interesting to see the relationship of free sulfur dioxide and chloride and sugar for each type of wine. 


## 1.7.1 Total Sulfur Dioxide: 
```{r total.sulfur.dioxide}
ggplot(wine, aes(total.sulfur.dioxide, fill=type))+
  geom_histogram(binwidth = 25)+
  scale_x_continuous(breaks = seq(0, 400, 25))+
  scale_y_sqrt()

```

```{r total.sulfur.dioxide summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(total.sulfur.dioxide), 
            median = median(total.sulfur.dioxide), 
            n = n())
```
>1.7.2 White wine also has a very high concentration of totla sulfur dioxide compared to the red wine. 


## 1.8.1 Density: 
```{r density}
ggplot(wine, aes(density, fill=type))+
  geom_histogram(binwidth = 0.001)+
  scale_x_continuous(breaks = seq(0, 2, 0.009))+
  scale_y_sqrt()
```

```{r density summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(density), 
            median = median(density), 
            n = n())
```
>1.8.2 Density of water is very similar for both types of wines, except there are few outliers that have almost two times more water density than the mean of the white wines. It might related to the outlier in the residual sugar variable. Literature states that water density is closely related with alcohol and residual sugar. We want to see if we can see any outliers in the alcohol variable too. I also want to test this later in our bivariate sections to see the relation of density , alcohol and residual sugar.


## 1.9.1 pH: 
```{r pH}
ggplot(wine, aes(pH, fill=type))+
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 10, 0.2))+
  scale_y_sqrt()
```

```{r pH summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(pH), 
            median = median(pH), 
            n = n())
```
>1.9.2 Both wines have a normal distribution for the the pH level. However it looks like teh red wine has a slightly higher pH than the white wine. pH is related to acidity of the wine, so it show that the red wine is more acidy. 

## 1.10.1 sulphates: 
```{r sulphates}
ggplot(wine, aes(sulphates, fill=type))+
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 3, 0.2))+
  scale_y_sqrt()
```

```{r sulphates summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(sulphates), 
            median = median(sulphates), 
            n = n())
```

>1.10.2 red wine has a higer sulphates amount than the white wine. Red wine has some outliers with sulphates more than 1.2 grams. 



## 1.11.1 Alcohol: 
```{r alcohol}
ggplot(wine, aes(alcohol, fill=type))+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(breaks = seq(0, 16, 0.5))+
  scale_y_sqrt()
```

```{r alcohol summary}
wine%>%
  group_by(type)%>%
  summarise(mean = mean(alcohol), 
            median = median(alcohol), 
            n = n())
```
>1.11.2 Both wine types have very similar alcohol percentage by volume, however white wines have a higher alcohol concentration. The above plot also shows that both wines peak at 9.5 % alcohol. 


## 1.12.1 Quality: 
```{r quality}
ggplot(wine, aes(quality, fill=type))+
  geom_bar()+
  scale_x_discrete(breaks = seq(0,10, 1))+
  scale_y_sqrt()
```

>1.12.2 The most frequent quality for both types of wine is 6 and then 5. It will be useful to see if a certain chemcial affects the wine quality. 




```{r alcohol}
 ggplot(aes(x=alcohol), data = wine)+
  geom_density(aes(fill=type))+
  facet_wrap(~quality)
names(r_wine)
```

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

The datasets that I chose for this analysis are very tidy and do not need a lot of wrangling. I had to combine both into one data set as "wine", since both datasets had the same variables. The combined dataset had a total of 6497 observations and 14 variables. 

Of the 14 variables, the first one is just the row number and the last variable indicates whether a particular observation is for red or white wine. 

### What is/are the main feature(s) of interest in your dataset?

I am intersted to find if there are certain chemicals that affect the quality of wine. Furthuremore, I want to know if there is a significant difference between the contents of red wine and white wine. If there is a difference, what are those and are those differnece contributing significantlly to the quality of the wine. 

In addition, I want to see if there is a significant correlation between certian components of wine. I am mainly interested to see the correlcation between chloride and residual sugar, alcohol and residual sugar, sulphates and residual sugar, ......

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

It will be helpful to look into sugar residual, chlorides, pH , alcohol volume, density and quality of each wine. 

### Did you create any new variables from existing variables in the dataset?

>I created one variable "type" to specify whether a wine is red or white. It was necessary for analysis of both types at once. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

> I noticed that some there are some wines with abnormal residual sugar, free sulfur dioxide, volatile acidity, and citric acid. 

#2. Bivariate Plots Section

> First I would like to plot a ggpairs plot of the only 1000 random sample of the wine against each other to see the relationship among the variables. 

```{r ggpairs}
#We will set the see so that our results are the same every time we run the code. 
set.seed(1122)
theme_set(theme_minimal(20))
#Plotting only 1000 of the observations: 
ggpairs(wine[sample.int(nrow(wine), 1000), ])
```

> Next, lets look at the correlation of the variables and see how correlated our variables are the wine quality:

```{r}
cor(x=wine[,2:12], y=wine$quality)
```

> Alcohol, density, chlorides and volatile acidity are correlated with wine quality. In the following section, we will explore these correlations in more details. 

## 2.1.1 Alcohol Vs. Quality
```{r alcohol vs. quality}

ggplot(aes(quality,alcohol, color=type), data = wine)+
  geom_point(position = 'jitter', alpha=0.2)+
  geom_smooth(method = 'lm', color='black')+
  facet_wrap(~type)
```
> We could see that higher quality wines tend to have more alcohol content. We could also see that the line for white wines is more steep than the red wine. This means the correlation of white wine with higher quailty is higher for white wines. 

## 2.2.1 Density vs. Quality
```{r density vs. quality}
ggplot(aes(quality,density, color=type), data = wine)+
  geom_point(position = 'jitter', alpha=0.5)+
  geom_smooth(method = 'lm', color='black')+
  facet_wrap(~type)
```
> The above plot shows that ther is a negative correlation between density and wine quality.. As the water density increases, the wine quality decreases. Again, it looks like the the correlation is stronger for white wines compared to red wines. 

## 2.3.1 Chlorides vs. Quality
```{r chlorides vs. quality}
ggplot(aes(quality,chlorides, color=type), data = wine)+
  geom_point(position = 'jitter', alpha=0.5)+
  geom_smooth(method = 'lm', color='black')+
  facet_wrap(~type)
```
>The above plot shows that there is negative and weak correlation between the chlorides and the wine qaulity

## 2.4.1 Chlorides vs. Quality
```{r volatile acidity vs. quality}
ggplot(aes(quality,volatile.acidity, color=type), data = wine)+
  geom_point(position = 'jitter', alpha=0.5)+
  geom_smooth(method = 'lm', color='black')+
  facet_wrap(~type)

```
> The above plot shows that there is a strong negative correlation between the volatile acidity of red wine and it's quality, however, the correlation of volatile and white wine is somehow weak. 





> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

## 2.1.1


```{r echo=FALSE, Bivariate_Plots}
names(wine)
ggplot(aes(x=pH, y=quality), data = wine)+
  geom_line(aes(color=type))

```

#  Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

## Alcohol and Quality relationship: 
```{r}
ggplot(aes(x=chlorides), data = wine)+
  geom_density(aes(fill=type))+
  scale_x_continuous(breaks = seq(0, 1.5, 0.07), lim = c(0, quantile(wine$chlorides, 0.99)))+
  facet_wrap(~quality)

```


```{r Chlorides vs. residual sugar}
ggplot(aes(chlorides, residual.sugar), data = wine)+
  geom_point(aes(color=type))+
  xlim(0, quantile(wine$chlorides, 0.99))+
  ylim(0, quantile(wine$residual.sugar, 0.99))+
  geom_smooth(method = 'lm')
names(wine)
```


```{r}
ggplot(aes(quality,pH), data = wine)+
  geom_point(alpha = 0.25, position = position_jitter(h=0), size = 4) +
  geom_smooth(method = "lm") +
  facet_wrap(~type)

ggplot(aes(y=alcohol, x=quality), data=wine) +
  geom_violin(aes(fill=quality), alpha=0.7) +
  ggtitle("Quality vs Alcohol Content") +
  xlab("Quality") +
  ylab("Alcohol (%)")
```

```{r}
ggplot(aes(y=residual.sugar, x=alcohol), data=wine) +
  geom_point(alpha=0.1) +
  geom_smooth(method='lm', formula=y~x) +
  ggtitle("Residual Sugar vs Alcohol Content") +
  xlab("Alcohol Content (%)") +
  ylab("Residual Sugar (g/L)") +
  ylim(0, 25)
```

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!

##Reference: 
  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.